Local Spectral Clustering of Density Upper Level Sets

### 2019-01-23 icml/icml_local_density_clustering.tex
Fixed few typos

\begin{theorem}
\label{thm: inverse_mixing_time_lower_bound_nonconvex}
Fix $\lambda > 0$, and let $\Cset \in \Cbb_f(\lambda)$ satisfy
Assumptions \ref{asmp: bounded_density} and \ref{asmp: cluster_diameter} 
for some $\sigma, \lambda_{\sigma}, \Lambda_{\sigma}, D > 0$. Then for any $0 <
r < \sigma/(4d)$, $0 < \delta < 1$, $\epsilon > 0$, and $n$ satisfying 
$$
\sqrt{3^{d+1}\frac{(\log n + d\log \mu + \log(4/\delta))}{n \nu_d r^d
    \lambda_{\sigma}}} \leq \epsilon, 
$$
where \smash{$\mu = \log(\frac{2D}{r})$}, we have
\begin{multline}
\label{eqn: inverse_mixing_time_lower_bound_nonconvex}
\Psi_{n,r}(\Csig[\Xbf]) \geq \left(d \log \mu + c_{\lambda} \right)^{-1} \cdot \\
\left(c_1 + c_2 \frac{\Lambda_{\sigma}^4D^{2d}}{\lambda_{\sigma}^4r^{2d}}\left(d
    \log \mu + c_{\lambda}\right)\right)^{-1},
\end{multline}
% RJT: confused ... is this correct?
with probability at least $1 - \delta$, where $c_1,c_2,c_3>0$ are constants
depending only on the dimension $d$, and $c_{\lambda} = \log(\Lambda_{\sigma}^2/
\lambda_{\sigma}^2)$.   
\end{theorem}
% RJT: saying that c_1,c_2,c_3 depend on the dimension is a little odd/weak.
% c_2 multiplyies, e.g., a factor of d \log \mu.  So if it's d^{10}, then this
% will swamp the linear factor of ....

\begin{remark}
The proof of Theorem \ref{thm: inverse_mixing_time_lower_bound_nonconvex} relies
on upper bounding the mixing time using the \emph{conductance} of $G_{n,r}[\Csig[X]]$, 
$$
\widetilde{\Phi} = \min_{S \subseteq \Csig[\Xbf]} \Phi(S; G_{n,r}[\Csig[X]]).
$$
The factor $1/r^{2d}$ in the lower bound in \eqref{eqn:
  inverse_mixing_time_lower_bound_nonconvex} is suboptimal. This exponential
dependence on $d$ stems from a loose bound on the aforementioned
conductance. In particular, we assert only that any set $S \subseteq \Csig[X]$
must have $\cut(S; \Csig[\Xbf])$ on the order of $n^2 r^{2d}$, while upper bounding
$\vol(S; \Csig[\Xbf])$ by roughly $n^2 r^d$, for a bound on the conductance of
order $r^d$. The presence of $r^{2d}$ comes from upper bounding the mixing time
by about \smash{$1/\widetilde{\Phi}^2$}, this being a variant of classic results
on rapid mixing \citep{jerrum89}.  
\end{remark}

@@

It is possible to sharpen the dependency on $d$ in 
\label{eqn: inverse_mixing_time_lower_bound_nonconvex}, but at the cost of an 
additional assumption: 
\begin{enumerate}[label=(A\arabic*)]
  \setcounter{enumi}{4}
\item 
  \label{asmp: convex}
  $\Cset$ is convex.
\end{enumerate}
% RJT: do we need C to satisfy this, or Csig?  Not sure I know why the former
% matters ...

With \ref{asmp: convex} in place, we now give a tighter lower bound on the
inverse mixing time of $C_\sigma[\Xbf]$ (albeit one that holds only
asymptotically).

\begin{theorem}
\label{thm: inverse_mixing_time_lower_bound}
Assume the conditions of Theorem \ref{thm:
  inverse_mixing_time_lower_bound_nonconvex}, and additionally, let $\Cset$
satisfy \ref{asmp: convex}. Then, 
for any $r < \sigma/4d$, the following statement holds with probability 1: 
\begin{align}
\label{eqn: inverse_mixing_time_lower_bound}
  \liminf_{n \to \infty} \, &\Psi_{n,r}(\Csig[\Xbf]) \geq \nonumber \\ 
&\quad c_1 \frac{\Lambda_{\sigma}^8}{\lambda_{\sigma}^8} (\log \mu +
  c_{\lambda}) \biggl( c_{\lambda}(d^3 \log\mu + c_2) +{} \nonumber \\  
&\quad \frac{d^2D^2}{r^2} \bigl( \log d + c_2
  \frac{\Lambda_{\sigma}^{2}}{\lambda_{\sigma}^{2}}c_{\lambda} +{} \nonumber \\  
&\quad c_3  + \log \log \mu \bigr) \biggr) + c_d \frac{o_r(1)}{r^2},
\end{align}
% RJT: confused ... is this correct?
where \smash{$\mu = \log(\frac{2D}{r})$}, $c_1,c_2,c_3>0$ are universal
constants, $c_d$ is constant in $r$ but depends on dimension $d$, and
$c_{\lambda} = \log(\Lambda_{\sigma}^2/ \lambda_{\sigma}^2)$.  
% Additionally, $\lim_{r \to 0} o_r(1) = 0$.  
\end{theorem}
% RJT: this was a lower bound on 1/\Psi and I changed it to \Psi ... just
% checking

@@

Our proof relies on a novel (to the best of our knowledge) uniform lower bound
on this conductance function over $G_{n,r}$ in terms of a population-level
analogue, which we term \smash{$\widetilde{\Phi}_{\Pbb,r}$}. Plugging
\smash{$\widetilde{\Phi}_{\Pbb,r}$} into \eqref{eqn: average_conductance} yields
a bound on mixing time (with respect to total variation distance) of order $d
D^2/r^2 + d^2\log(D/r)$ over convex sets. By contrast, \eqref{eqn:
  inverse_mixing_time_lower_bound} is of order $d^2 D^2/r^2 + d^3\log(D/r)$ 
(ignoring the $o_r(1)$ term) but handles mixing time with respect to relative
pointwise distance (which is known to be a stricter metric). 
\end{remark}

\begin{remark}
  The convexity requirement is necessary only to lower bound the
  \smash{$\widetilde{\Phi}_{\Pbb,r}$}; any bound on $\widetilde{\Phi}_{\Pbb,r}$
  which   does not require convexity could immediately be plugged into the
  machinery of   the proof of Theorem \ref{thm: inverse_mixing_time_lower_bound}
  to achieve a corresponding result. Recent work \citep{abbasi17} has developed
  such population-level bounds on the conductance function, however the Markov
  chain dealt with there is somewhat different than the one we consider.  
\end{remark}

>>>>

\begin{theorem}
\label{thm: inverse_mixing_time_lower_bound_nonconvex}
Fix $\lambda > 0$, and let $\Cset \in \Cbb_f(\lambda)$ satisfy
Assumptions \ref{asmp: bounded_density} and \ref{asmp: cluster_diameter} 
for some $\sigma, \lambda_{\sigma}, \Lambda_{\sigma}, D > 0$. Then for any $0 <
r < \sigma/(4d)$, $0 < \delta < 1$, $\epsilon > 0$, and $n$ satisfying 
$$
\sqrt{3^{d+1}\frac{(\log n + d\log \mu + \log(4/\delta))}{n \nu_d r^d
    \lambda_{\sigma}}} \leq \epsilon, 
$$
where \smash{$\mu = \log(\frac{2D}{r})$}, we have
\begin{multline}
\label{eqn: inverse_mixing_time_lower_bound_nonconvex}
\Psi_{n,r}(\Csig[\Xbf])^{-1} \leq (d \log \mu + c_{\lambda}) \cdot \\  
\left(c_1 + c_2 \frac{\Lambda_{\sigma}^4D^{2d}}{\lambda_{\sigma}^4r^{2d}}\left(d
    \log \mu + c_{\lambda}\right)\right),
\end{multline}
with probability at least $1 - \delta$, where $c_1,c_2,c_3>0$ are constants
depending only on the dimension $d$, and $c_{\lambda} = \log(\Lambda_{\sigma}^2/
\lambda_{\sigma}^2)$.   
\end{theorem}
% RJT: saying that c_1,c_2,c_3 depend on the dimension is a little odd/weak.
% c_2 multiplyies, e.g., a factor of d \log \mu.  So if it's d^{10}, then this
% will swamp the linear factor of ....

\begin{remark}
The proof of Theorem \ref{thm: inverse_mixing_time_lower_bound_nonconvex} relies
on upper bounding the mixing time using the \emph{conductance} of $G_{n,r}[\Csig[X]]$, 
$$
\widetilde{\Phi} = \min_{S \subseteq \Csig[\Xbf]} \Phi(S; G_{n,r}[\Csig[X]]).
$$
The factor of $1/r^{2d}$ in the bound in \eqref{eqn:
  inverse_mixing_time_lower_bound_nonconvex} is suboptimal. This exponential 
dependence on $d$ stems from a loose bound on the aforementioned
conductance. In particular, we assert only that any set $S \subseteq \Csig[X]$
must have $\cut(S; \Csig[\Xbf])$ on the order of $n^2 r^{2d}$, while upper bounding
$\vol(S; \Csig[\Xbf])$ by roughly $n^2 r^d$, for a bound on the conductance of
order $r^d$. The presence of $r^{2d}$ comes from upper bounding the mixing time
by about \smash{$1/\widetilde{\Phi}^2$}, this being a variant of classic results
on rapid mixing \citep{jerrum89}.  
\end{remark}

@@

It is possible to sharpen the dependency on $d$ in 
\eqref{eqn: inverse_mixing_time_lower_bound_nonconvex}, but at the cost of an  
additional assumption.
\begin{enumerate}[label=(A\arabic*)]
\setcounter{enumi}{4}
\item 
\label{asmp: convex}
\emph{Convexity:} The set $\Cset_\sigma$ is convex. 
\end{enumerate}
% RJT: do we need C to satisfy this, or Csig?  Not sure I know why the former
% matters ...

With \ref{asmp: convex} in place, we give a tighter bound on the
mixing time (albeit one that holds only asymptotically).

\begin{theorem}
\label{thm: inverse_mixing_time_lower_bound}
Assume the conditions of Theorem \ref{thm:
  inverse_mixing_time_lower_bound_nonconvex}, and additionally, \ref{asmp:
  convex}. Then for any $r < \sigma/4d$, the following holds with probability 1:  
\begin{align}
\label{eqn: inverse_mixing_time_lower_bound}
  \limsup_{n \to \infty} \, &\Psi_{n,r}(\Csig[\Xbf])^{-1} \leq \nonumber \\ 
&\quad c_1 \frac{\Lambda_{\sigma}^8}{\lambda_{\sigma}^8} (\log \mu +
  c_{\lambda}) \biggl( c_{\lambda}(d^3 \log\mu + c_2) +{} \nonumber \\  
&\quad \frac{d^2D^2}{r^2} \bigl( \log d + c_2
  \frac{\Lambda_{\sigma}^{2}}{\lambda_{\sigma}^{2}}c_{\lambda} +{} \nonumber \\  
&\quad c_3  + \log \log \mu \bigr) \biggr) + c_d \frac{o_r(1)}{r^2},
\end{align}
where \smash{$\mu = \log(\frac{2D}{r})$}, $c_1,c_2,c_3>0$ are universal
constants, $c_d$ is constant in $r$ but depends on dimension $d$, and
$c_{\lambda} = \log(\Lambda_{\sigma}^2/ \lambda_{\sigma}^2)$.  
% Additionally, $\lim_{r \to 0} o_r(1) = 0$.  
\end{theorem}
% RJT: this was a lower bound on 1/\Psi and I changed it to \Psi ... just
% checking

@@

Our proof relies on a novel (to the best of our knowledge) uniform lower bound
on this conductance function over $G_{n,r}$ in terms of a population-level
analogue, which we denote \smash{$\widetilde{\Phi}_{\Pbb,r}$}. Plugging
\smash{$\widetilde{\Phi}_{\Pbb,r}$} into \eqref{eqn: average_conductance} yields
a mixing time bound (with respect to total variation distance) of order $d
D^2/r^2 + d^2\log(D/r)$ over convex sets. By contrast, \eqref{eqn:
  inverse_mixing_time_lower_bound} is of order $d^2 D^2/r^2 + d^3\log(D/r)$ 
(ignoring the $o_r(1)$ term) but handles mixing time with respect to relative
pointwise distance (which is known to be a stricter metric). 
\end{remark}

\begin{remark}
  The convexity requirement is necessary only to lower bound the
  \smash{$\widetilde{\Phi}_{\Pbb,r}$}. Any bound on
  \smash{$\widetilde{\Phi}_{\Pbb,r}$} which does not require convexity could
  immediately be plugged into the machinery of the proof of Theorem \ref{thm:
    inverse_mixing_time_lower_bound} to achieve a corresponding result. Recent
  work \citep{abbasi17} has developed such population-level bounds on the
  conductance function, however the Markov chain dealt with there is somewhat
  different than the one we consider.   
\end{remark}
